---
title: "Flight Data Analysis"
author: "Frederick Jones"
date: "October 1, 2023."
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
   warning=FALSE, message=FALSE,  tidy=TRUE,	tidy.opts=list(arrow=TRUE,	indent=2),fig.width = 10, fig.align = 'center', fig.asp=0.618, out.width = '100%')
```
# Load the required libraries
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(flextable)
library(rstatix)
```

# GitHub URL for the CSV file
```{r}
url <- "https://raw.githubusercontent.com/jewelercart/Data607/main/flight_data.csv"
```

# Read the CSV file into R, specifying header and line termination
```{r}
flight_data <- read.csv(url, header = TRUE, sep = ",", quote = "", fill = TRUE)

```

# Specify the column names explicitly
```{r}
colnames(flight_data) <- c("Time_Zone", "Cities", "on_time", "delayed")

```

# Tidy the data (convert to long format)
```{r}
flight_data_long <- flight_data %>%
  gather(key = "Status", value = "Count", -Time_Zone, -Cities)
```


# Removing the quotes from the names of Time Zones and Cities.
```{r}
flight_data_long$Time_Zone<- gsub("\"", "", flight_data_long$Time_Zone)
flight_data_long$Cities<- gsub("\"", "", flight_data_long$Cities)
table <- knitr::kable(flight_data_long)
table
```



# Calculate summary statistics, handling missing values
```{r}
summary_stats =
flight_data_long  %>%
  group_by(Time_Zone, Status) %>%
   get_summary_stats(Count, show = c("mean", "median", "max","min")) 
```

# Create bar plots to compare arrival delays
```{r}
ggplot(flight_data_long, aes(x = Time_Zone, y = Count, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Comparison of Arrival Delays by Time Zones and Status", y = "Number of Flights") +
  theme_minimal()
```

# Filter the data for ALASKA and AM WEST separately:
```{r}
alaska_data <- flight_data_long %>% filter(Time_Zone == "ALASKA")
am_west_data <- flight_data_long %>% filter(Time_Zone == "AM WEST")
```

# Create bar plots for ALASKA and AM WEST:
```{r}
plot_alaska <- ggplot(alaska_data, aes(x = Cities, y = Count, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Arrival Delays for ALASKA by City and Status", y = "Number of Flights") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust=0.5),
        panel.grid.minor= element_blank())

plot_alaska
```


```{r}
plot_am_west <- ggplot(am_west_data, aes(x = Cities, y = Count, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Arrival Delays for AM WEST by City and Status", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust=0.5),
        panel.grid.minor= element_blank())

plot_am_west
```


Summary Statistics:
```{r}
summary_stats
```

# A Chi-square test will compare whether the delay difference of a flight varies by comparing two time zones.
```{r}
tbl <- xtabs(~ Time_Zone + Status, data = flight_data_long)
summary(tbl)
```

```{r}
proportions(tbl, "Status")
```
# Conclusion

There is no difference between the two time zones regarding flight delays and punctuality. This was confirmed with a chi-squared test p-value = 1.
