---
title: "Week 3 Assignment"
author: "Frederick Jones"
date: "2023-09-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
   warning=FALSE, message=FALSE,  tidy=TRUE,	tidy.opts=list(arrow=TRUE,	indent=2),fig.width = 10, fig.align = 'center', fig.asp=0.618, out.width = '100%')
```

# Libraries
```{r}
library(tidyverse)
library(here)
```

# First Question 

```{r}
data = read.csv(here('majors-list.csv'))

data |> select('Major') |> filter(str_detect(Major, 'DATA|STATISTICS'))
```

# Second Question

Data:
[1] "bell pepper" "bilberry" "blackberry" "blood orange"
[5] "blueberry" "cantaloupe" "chili pepper" "cloudberry"
[9] "elderberry" "lime" "lychee" "mulberry"
[13] "olive" "salal berry"

```{r}
list = c("bell pepper", "bilberry", "blackberry", "blood orange","blueberry", "cantaloupe", "chili pepper", "cloudberry","elderberry", "lime", "lychee", "mulberry","olive", "salal berry")
list

```

# Third Question:  Describe, in words, what these expressions will match:

**(.)\1\1 =>** This regular expression identifies sequences of three consecutive identical characters.` 

```{r}
str_detect(c( "AAA", "999","99"),"(.)\\1\\1")
```


**(.)(.)\\2\\1 =>** This regular expression detects strings consisting of four characters,
where the first and last characters are identical, while the two middle characters are also identical but distinct from the first and last.  
```{r}
str_detect(c("noon","1212","xyyx"),'(.)(.)\\2\\1')
```




**(..)\1 =>** This regular expression pinpoints strings of at least four characters, where the initial two characters are equivalent to the subsequent two characters.` 
```{r}
str_detect(c('noon',"1212","xyyx"),'(..)\1')
```



**(.).\\1.\\1 =>** This regular expression recognizes strings comprising five characters, where the first character matches the third and fifth characters, and the second character differs.` 

```{r}
str_detect(c("AXA","121","bYb"),'(.).\\1.\\1')
```

**(.)(.)(.).*\\3\\2\\1 =>** This regular expression identifies strings of at least six characters, where the first three characters form a palindrome.`
```{r}
str_detect(c("121AABBA", "xyx12345zyx","wow987654321"),'(.)(.)(.).*\\3\\2\\1')
```

# Forth Question: Construct regular expressions to match words that

**Start and end with the same character.**
```{r}
str_detect('tarot','^(.).*\\1$')
```

**Contain a repeated pair of letters.**
```{r}
str_detect("church",".*(.{2}).*\\1.*")

```

**Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)**
```{r}
str_detect("eleven","([a-z]).*\\1.*\\1")
```

